---
title: "general_netcdf_process"
author: "fx"
date: "2020年12月8日"
output: html_document
---

# 1 get the basic infromation

- variables and attributes
```{r}
library(ncdf4)
filelist <- list.files(path="",pattern="A2006") 
ncfilename <- filelist[1]

# open ncfile
ncfile <- nc_open(ncfilename)   

print(ncfile) # detail information of the file
# get variable
lon <- ncvar_get(ncfile,"lon")
lat <- ncvar_get(ncfile,"lat")
tmp <- ncvar_get(ncfile,"tmp")

# get attributes
tlongname <- ncatt_get(ncfile,"tmp","long_name")
tmp_units <- ncatt_get(ncfile,"tmp","units")
tmp_fillvalue <- ncatt_get(ncfile,"tmp","_FillValue")
timeunit <- ncatt_get (ncfile,"time","units") # name and attributes, the time since
title <- ncatt_get(ncfile,0,"title") ## get whole attributes
```

# 2 reshape 

## convert time variable
-if there is no information about the current time ,conversion of time variable is needed, ** timeunit$value(time-since)-->real time**

```{r}
library(lattice)
library(RColorBrewer)
library(chron)

tstr <- strsplit(timeunit$value,"")
tdstr <- strsplit(unlist(tstr)[3],"-")
tday <- as.integer(unlist(tdstr))
chron()
```

## NA exchange
```{r}
tmp[tmp==tmp_fillvalue$value] <- NA
```

## single slice transform
```{r}
# the tmp obtained before is a muti-dimention ,muti layers
tmp_slice <- tmp[,,1]

# quick map
image(lon,lat,tmp_slice,col = rev(brewer.pal(10,"RdBu")))

# better map
tmp_grid <- expand.grid(lon=lon,lat=lat)
cutpts <- seq(from=-50,to=50,by=10)
levelplot(tmp_slice~lon*lat,data = tmp_grid,at=cutpts,cuts=11,pretty=TRUE,col.region=rev(brewer.pal(10,"RdBu")))

# transform into a dataframe and save as a CSV file
lonlat <- as.matrix(expand.grid(lon,lat)) # grid --> matrix
tem_vector <- as.vector(tmp_slice)  # slice(grid)--->vector
tmp_df <- data.frame(cbind(lonlat,tmp_vector))
names(tmp_df) <- c("lon","lat","tmpday1")
write.table(tmp_df,"tmp_df.csv",row.names = FALSE) # 以往存储的时候没有后面的rownames=False参数，导致多了一列，以后要注意哦

```
## whole array transform(include time dimention)
```{r}
tmp_vec_long <- as.vector(tmp) # many grids -->a long vector
tmp_matrix <- matrix(tmp_vec_long,nrow =nlon*nlat,ncol=nt) # vector---> matrix(nlon*nlat * nt(time) )
tmp_df_wider <- data.frame(cbind(lonlat,tmp_matrix))
names(tmp_df_wider) <- c("lon","lat","t1",paste0("t",seq(from=1,to=12,by=1)))
# calculation based on df file
tmp_df_wider$annualmean <- apply(tmp_df_wider[3:14],1,mean) # Note the right number3+12-1
```
# 3 dataframe into array

## a full dataframe into array
```{r}
# get lon,lat,time, nlon,nlat,ntime
# rechange
tmp_matrix2 <- as.matrix(tmp_df_wider[3:3+12-1])  # df--->matrix
tmp_arry2 <- array(tmp_matrix2,dim=c(nlon,nlat,nt))
mean_arry2 <- array(tmp_df_wider$annualmean,dim=c(nlon,nlat)) # 单独的一列统计值，如何处理？
# check
levelplot(tmp_arry2[,,1]~lon*lat,data=grid,at=cutpts,cut=11,pretty=T,col.regions=rev(brewer.pal(10,"RdBu")),title="January") # data= Grid??
```
## a short dataframe into array

- loop
```{r}
fillvalue <- 1e32
tmp_arry3 <- array(fillvalue,dim = c(nlon,nlat,nt))
mean_arry3 <- array(fillvalue,dim = c(nlon,nlat))
# copy 
tmp_df_wider3 <- na.omit(tmp_df_wider)
nobs <- dim(tmp_df_wider3)[1] #12
for (i in 1:nobs) {
  j <- which.min(abs(lon-tmp_df_wider3$lon[i]))
  k <- which.min(abs(lat-tmp_df_wider3$lat[i]))
  
  tmp_arry3[j,k,1:nt] <- as.matrix(tmp_df_wider3[i,3:(nt+2)])
  mean_arry3[j,k] <- as.matrix(tmp_df_wider3$annualmean[i])
}
```

- loop(simple example)
```{r}
lon <- c(1,2,3,4,5)
lat <- c(3,4,5,6,7)
df <- as.data.frame( expand.grid(lon=lon,lat=lat))
df$sample <- sample(25)
df$sample2 <- sample(25)
df_array <- array(fillvalue,dim = c(5,5,2)) 
nobs <- dim(df)[1] # 行数25


for (i in 1:nobs) { #一列一列地遍历
  j <- which.min(  abs(lon-df$lon[i]))  # array的位置，j纵坐标，K横坐标
  k <- which.min(  abs(lat-df$lat[i]))
  df_array[j,k,1:2] <- as.matrix(df[i,3:4])  # 按列存放的性质
  
}
dim(df_array )
```
- semi-loop
```{r}
j2 <- sapply(df$lon, function(x) which.min(abs(lon-x))) # 所在的lon开始位置
k2 <- sapply(df$lat, function(x) which.min(abs(lat-x)))
j2
k2
df_array1 <- array(fillvalue,dim = c(5,5))
for (i in 1:2) {
  df_array1[cbind(j2,k2)] <- as.matrix(df[,2+i])
  df_array[,,i] <-df_array1
}
df_array

```
-no-loop
```{r}
nt<-2
l <- rep(1:nt,each=nobs)
df_array[cbind(j2,k2,l)]<-as.matrix(df[,(3:3+nt-1)]) # 厉害  
df_array
m <- 1
levelplot()
levelplot(df_array[,,m]~lon*lat,cuts = 11,at=cutpts,col.regions=rev(brewer.pal(10,"RdBu")))
library(RColorBrewer)
library(lattice)
```
# 4 creat netcdf file
```{r}
# 创建文件名：
ncname <- "new.nc"
dname <- "tmp"
nt<-2
# 定义内容
londim <- ncdim_def("lon","degree_east",as.double(lon))

latdim <- ncdim_def("lat","degree_north",as.double(lat))

timedim <- ncdim_def("time","times",as.double(nt))

fillvalue <- 1e32

tmp_def <- ncvar_def("tmp","C",list(londim,latdim,timedim),fillvalue,prec = "single")
ncout <- nc_create(ncname,list(tmp_def),force_v4 = TRUE)
ncvar_put(ncout,tmp_def,df_array)
ncatt_put(ncout,"lon","axis","X")
ncatt_put(ncout,"lat","axis","Y",)
ncatt_put(ncout,"time","axis",'T')
nc_close(ncout)

```

# 5 reading and writting  projected file
```{r}
# get coordinate variable 
x <- ncvar_get(ncfile,"x")
y <- ncvar_get(ncfile,"y")
# get projection
grid_mapping_name <- ncatt_get(ncfile,"Lambert_Conforma1","grid_mapping_name")
standard_parallel <- ncatt_get(ncfile,"Lambert_Conforma1","standard_parallel")
```



